---
title: "Draft of the proposed sampling framework"
format: html
execute:
  echo: false
---


```{r}
#| label: setup
library(here)
library(knitr)
library(terra)
library(tidyverse)
dominant <- here("data", "dominant")
type <- "niv1-estat"
version <- "v1-2"
year <- 2022
```

- weighted fraction of every level 1 class per pixel
  - in a circle with radius of 200 m around the pixel
  - bivariate gaussian kernel, sigma = 7.77
      - 99% of the density is within 200 m
      - larger weights for closer pixels, smaller weights for more distant pixels
- lump `heather` and `sparse vegetation` into a single `heather` class
- lump `river`, `lake` and `inlet` into a single `water` class
- determine the dominant class
  - `no minimum`: the dominant class is the one with the highest weighted fraction
  - `third`: the dominant class is the one with the highest weighted fraction, but only if it is at least 1/3 of the total weighted fraction
  - `half`: the dominant class is the one with the highest weighted fraction, but only if it is at least 1/2 of the total weighted fraction
  - `twothird`: the dominant class is the one with the highest weighted fraction, but only if it is at least 2/3 of the total weighted fraction
- `water` is handled differently, because it is a special case
  - `open water`: the dominant class is `water`, and the weighted fraction of `water` is greater than 1/3 of the total weighted fraction.
    These pixels are often not accessible for sampling.
  - `water edge`: the dominant class is `water`, but the weighted fraction of `water` is less than or equal to 1/3 of the total weighted fraction
- Assign the remaining pixels to the `heterogeneous` stratum, because they are not dominated by any class, according to the definition of the dominant class.

```{r}
#| label: read-map
sprintf("%s/dominant-strata-%s-%s-%i.tiff", dominant, type, version, year) |>
  rast() |>
  as.vector() |>
  table() -> tab
data.frame(
  code = names(tab),
  dominant = floor(as.integer(names(tab)) / 10),
  fraction = as.integer(names(tab)) - 10 * floor(as.integer(names(tab)) / 10),
  cells = as.vector(tab)
) |>
  mutate(
    no_minimum = case_match(
      .data$dominant,
      1 ~ "urban",
      2 ~ "cropland",
      3 ~ "grasland",
      4 ~ "forest",
      c(5, 6) ~ "heather",
      7 ~ "wetland",
      c(8, 9, 10) ~ "water",
      c(11) ~ "dune"
    ),
    no_minimum = ifelse(
      .data$no_minimum == "water",
      ifelse(.data$fraction <= 1, "water edge", "open water"),
      .data$no_minimum
    ),
    third = ifelse(
      .data$fraction < 1 & !grepl("water", .data$no_minimum),
      "heterogeneous",
      .data$no_minimum
    ),
    half = ifelse(
      .data$fraction < 2 & !grepl("water", .data$no_minimum),
      "heterogeneous",
      .data$no_minimum
    ),
    twothird = ifelse(
      .data$fraction < 3 & !grepl("water", .data$no_minimum),
      "heterogeneous",
      .data$no_minimum
    )
  ) -> strata
```


```{r}
#| label: tbl-strata
#| tbl-cap: "Number of pixels in each stratum, for different definitions of the
#|  dominant class.
#|  The `heterogeneous` stratum contains pixels where the dominant class is not
#|  dominant enough, according to the definition of the dominant class."
strata |>
  group_by(stratum = .data$no_minimum) |>
  summarise(cells = sum(.data$cells), type = "no minimum") |>
  bind_rows(
    data.frame(stratum = "heterogeneous", cells = 0, type = "no minimum"),
    strata |>
      group_by(stratum = .data$third) |>
      summarise(cells = sum(.data$cells), type = "third"),
    strata |>
      group_by(stratum = .data$half) |>
      summarise(cells = sum(.data$cells), type = "half"),
    strata |>
      group_by(stratum = .data$twothird) |>
      summarise(cells = sum(.data$cells), type = "twothird")
  ) |>
  mutate(
    cells = sprintf("%.2f%%", 100 * .data$cells / sum(strata$cells)),
    stratum = factor(.data$stratum) |>
      reorder(.data$cells, FUN = max, decreasing = TRUE)
  ) |>
  pivot_wider(names_from = "type", values_from = "cells") |>
  arrange(.data$stratum) |>
  kable()
```
